Fuzzy Extractors for Noisy Sources \\with More Errors Than Entropy

### 2015-02-11 trunk/constantRate/ConstCompFuzzExt-crypto-full.tex
adding version for crypto with appendices pulled forward

\subsubsection{Reusable Fuzzy Extractors}
\label{sec:reusable}

@@

We modify the definition of Boyen~\cite[Definition 6]{Boyen2004} for the computational setting.  We discuss the our definition and definitions due to Boyen in \apref{sec:comparing reusability}.

@@

\subsection{Obfuscation}
Our constructions will use obfuscation for two types of circuits: point functions and digital lockers. The family of point functions $\mathtt{I}_n = \{I_w\}_{w \in \zo^n}$ defined as follows:
\[
I_w(x):\begin{cases} 1 & x=w\\0 & \text{otherwise}\end{cases}.
\]
and the class of digital lockers is $\mathtt{I}_n = \{I_{w, r}\}_{w \in \zo^n, r\in\zo^\kappa}$ defined as follows:
\[
I_{w, r}(x):\begin{cases} r & x=w\\\perp & \text{otherwise}\end{cases}.
\]
The required notion of obfuscation is virtual grey-box (VGB) introduced in \cite{bitansky2010strong}. This notion is weaker then the standard notion of virtual black-box (\cite{barak2001possibility}), as it allows the simulator to run in unbounded time while making at a polynomial number of oracle queries to the function. VGB obfuscators for point functions and digital lockers are constructible from specific number-theoretic assumptions or by strong assumptions on hash functions.  We provide more details and a formal definition in \apref{sec:obfuscation def}.

@@

 Correctness follows immediately: for any given $i$, the probability that $v'_i=v_i$ is at least $1/p(n)$. By choosing $\ell$ sufficiently bigger than $1/p(n)$, we guarantee that for at least one $i$,  $v'_i=v_i$ with all but negligible probability
(indeed, because the $\ell$ samples are independent, the probability this does not happen is at most $(1-1/p(n))^\ell \le e^{-\ell/p(n)}$, using $1+x \le e^x$).

Security follows from obfuscation and is analyzed in \apref{sec:analysis sampling}.

@@

\subsection{Reusability of \consref{cons:sampling}}
The reusability of \consref{cons:sampling} follows from the security of the VGB obfuscator with auxiliary input. Consider any $q = \poly(n)$ number of reuses.  For each fixed $i\in \{1,..., q\}$, we can treat the keys $r_1,\dots, r_{i-1}, r_{i+1}, \dots, r_q$ and the strings $p_1, \dots, p_q$ as auxiliary input to the obfuscation adversary. The result follows by simulatability of this adversary, using the same argument as the proof of Theorem~\ref{thm:sampling} in \apref{sec:analysis sampling}. Note that the simulator in the definition of composable obfuscation is required to function for arbitrary circuits in the family even if the choice of these circuits depends on the previous obfuscations, which allows the $i$th reading of $w$ to be chosen depending on public values $p_1,..., p_{i-1}$, as required by definition of reusability.

@@

\noindent
The following theorem shows the security of this construction (proof is in \apref{sec:info theory sec}).

\begin{theorem}
\label{thm:info theory sec}
%Let $n$ be a security parameter and let $\gamma = \omega(\log n)$.
Let $\mathcal{W}$ be a family of $(\alpha = \Omega(1), \beta\leq \gamma(1-\Theta(1)))$-sparse block sources over $\mathcal{Z}^\gamma$ and let $\cond: \mathcal{Z} \times \zo^d\rightarrow \mathcal{Y}$ be a $(\alpha, \tilde{\alpha}, \epsilon_{cond})$-randomness conductor.  Define $\mathcal{V}$ as the family of all distributions with min-entropy at least $\tilde{\alpha}(\gamma-\beta)$ and let $(\gen', \rep')$ be $(\mathcal{Y}^\gamma, \mathcal{V}, \kappa, t, \epsilon_{fext})$-fuzzy extractor with error $\delta$.\footnote{We actually need $(\gen', \rep')$ to be an average case fuzzy extractor~(see \cite[Definition 4]{DBLP:journals/siamcomp/DodisORS08} and the accompanying discussion).  Most known constructions of fuzzy extractors are average-case fuzzy extractors.  For simplicity we refer to $\gen', \rep'$ as simply a fuzzy extractor.}  Then $(\gen, \rep)$ is a $(\mathcal{Z}^\gamma, \mathcal{W}, \kappa, t, \gamma\epsilon_{cond}+\epsilon_{fext})$-fuzzy extractor with error $\delta$.
\end{theorem}

@@

We first define a slightly weaker object than a computational fuzzy extractor: it outputs a key with computational entropy~(instead of a pseudorandom key).  
We call this object a computational fuzzy conductor.  It is the computational analogue of a fuzzy conductor~(introduced by Kanukurthi and Reyzin~\cite{KanukurthiR09}).
Before defining this object, we define conditional computational (``HILL'' \cite{DBLP:journals/siamcomp/HastadILL99}) entropy.

@@

\subsection{Reusability of Fuzzy Extractors}
\label{sec:comparing reusability}

The goal of a reusable fuzzy extractor is to allow enrollment of a source across multiple services.  A service $i$ sees an reading of the source $w_i$.
Boyen considers two versions of reusable fuzzy extractors, first where the adversary sees $p_1,..., p_q$~(outsider security~\cite[Definition 6]{Boyen2004}) and tries to learn about the values $w_1,..., w_q$ or the keys $r_1,..., r_q$.  Second, where the adversary controls some subset of the servers and can key generation on arbitrary $p_i'$~(insider security~\cite[Definition 7]{Boyen2004}).   This allows the adversary to learn a subset of keys $r_i$~(by performing key generation on the valid $p_i$).  This definition makes sense when servers are compromised~(after enrollment) and act maliciously.  In both definitions, the adversary creates a perturbation function $_i$ after seeing $p_1,..., p_{i-1}$~(and generated keys for outsider security) and the challenger generates $w_i = f_i(w_1)$.  The definition is parameterized by the class of allowed perturbation functions.

We adapt the definition of Boyen to the computational setting~(\defref{def:outsider fuzz ext}).  
The definition of Boyen considers a single adversary.  We split the adversary into two parts, one of which is information-theoretic and another that is computationally bounded.  The functions $f_2,..., f_q$ can be thought of as a single adversary that sees all prior state.  However, to provide meaningful security in the computational setting, we cannot have communication between $f_2,..., f_q$ and $D$.  This is because $f_2,..., f_q$ are computationally unbounded and thus could break any security that only holds against computationally bounded adversaries.
Because these two adversaries do not communicate we strengthen the definition by allowing the perturbation functions,  $f_i$, to see the original sample $w_1$.\footnote{An alternative would be to have a single computationally bounded adversary that provides perturbation functions but does not see the original reading $w_1$ as in the definition of Boyen.  \consref{cons:sampling} also satisfies this alternative adaption of Boyen's definition.}  This was not allowed in the definition of Boyen, because the adversary was not split-state, and so giving $w_1$ to the adversary would make security impossible.

%Finally, we allow the distinguishing adversary to see other keys \lnote{we do? then why do we call our definition ``outsider''?}.  This is strictly stronger than the definition of Boyen \lnote{you mean than the outsider security definition of Boyen?}.  However, it is weaker than the notion of insider reusable fuzzy extractors where the adversary sees key generation on arbitrary $p$.  This \lnote{what is ``this''?} is a mixture of reusability and robustness~\cite{dkrs2006} of the fuzzy extractor.  Our reusable construction~(\consref{cons:sampling}) is not insider secure \lnote{according to Boyen's notion, right? Why not say ``not robust'' and/or clarify that $r_i$ can leak but $p_i$ cannot be modified.}.

\subsection{Obfuscation}
\label{sec:obfuscation def}

In this section, we give a formal definition of the required notion of obfuscation.
We require that the obfuscation is composable and secure with respect to auxiliary input. Composable auxiliary-input VGB obfuscators for point functions and digital lockers are constructed in \cite[Theorem 6.1]{bitansky2010strong} (following \cite{canetti2008obfuscating,CKVW10}) from the Strong Vector Decision Diffie-Hellman assumption, which is a generalization of the strong DDH assumption of \cite{canetti1997towards} for tuples of points. As described in the Introduction, they can also be constructed by assuming strong properties of cryptographic hash functions~\cite{canetti1997towards}, \cite[Section 4]{lynn2004positive}, \cite[Section 3.2]{canetti2008obfuscating}, \cite[Section 8.2.3]{dakdoukThesis} (such a construction results in an obfuscated function that, at a few hard-to-find-inputs, differs from the original function; this change does not materially affect our results).

@@

\section{Proof of Security of \consref{cons:sampling}}
\label{sec:analysis sampling}
 Our goal is to show that for all $s_{sec} = \poly(n)$ there exists $\epsilon_{sec} =\ngl(n)$ such that $\delta^{\mathcal{D}_{s_{sec}}}((R, P), (U, P))\le \epsilon_{sec}$. Let $D$ be a distinguisher of size at most $s_{sec}$. We want to bound 
\[
| \expe[D(R, P)] - \expe[D(U, P)]|
\]
by a negligible function.

Adding together Equation~\ref{eq:dist before sampling}, Equation~\ref{eq:sim cannot distinguish samp}, and Equation~\ref{eq:dist before sampling} in which $R$ is replaced with $U$, we obtain the desired result by the triangle inequality.\section{Analysis of \consref{cons:first construction}}
\label{sec:construction analysis}

@@

\section{Analysis of \consref{cons:info theoretic}}
\label{sec:info theory sec}
\begin{proof}[Proof of Theorem~\ref{thm:info theory sec}]
Let $W\in \mathcal{W}$.
It suffices to argue correctness and security.  We first argue correctness.
When $w_i = w_i'$, then $\cond(w_i , seed_i) = \cond(w_i', seed_i)$ and thus $v_i = v_i'$.  Thus, for all $w, w'$ where $\dis(w, w')\le t$, then $\dis (v, v')\le t$.  Then by correctness of $(\gen', \rep')$, $\Pr[(r, p)\leftarrow \gen'(v) \wedge r'\leftarrow \rep(v',p) \wedge r' = r]\ge 1-\delta$.

We now argue security.  Denote by $seed$ the random variable consisting of all $\gamma$ seeds and $V$ the entire string of generated $V_1,..., V_\gamma$.  To show that $R | P, seed \approx_{\gamma \epsilon_{cond} + \epsilon_{fext}} U | P, seed$, it suffices to show that $\Hav(V | seed)$ is $\gamma \epsilon_{cond}$ close to a distribution with average min-entropy $\tilde{\alpha}(\gamma - \beta)$.  The lemma then follows by the security of $(\gen', \rep')$.\footnote{Note, again, that $(\gen', \rep')$ must be an average-case fuzzy extractor.  Most known constructions are average-case and we omit this notation.}

\noindent
By a hybrid argument there exists a distribution $Z$ with $\Hav(Z | seed) \ge \tilde{\alpha}(\gamma -\beta)$ where
\[
(\cond(W_\gamma, seed_\gamma), ..., \cond(W_1, seed_1), seed_\gamma,..., seed_1) \approx_{\gamma \epsilon_{cond}} (Z, seed_\gamma,...,  seed_1).\]
%By the security of $(\sketch, \rec)$ we know that $\Hav(Z | seed, ss) \ge \tilde{m}$.  %Note that $U_{\mathcal{Y}}^\gamma$ is independent of $seed_1,..., seed_\gamma$ and so $\Hav(U_{\mathcal{Y}}^\gamma, seed_1,..., seed_\gamma, p)\ge \tilde{m}$.
%That is,
%\[
%(V, seed_1,..., seed_\gamma, p) \approx_{\epsilon\gamma} (U_{\mathcal{Y}}^\gamma, seed_1,..., seed_\gamma, p).\]
This completes the proof.
\end{proof}

>>>>

\subsection{Reusable Fuzzy Extractors}
\label{sec:reusable}

@@

We modify the definition of Boyen~\cite[Definition 6]{Boyen2004} for the computational setting.  We first present our definition and then compare to the definitions of Boyen.

@@

\paragraph{Comparison with the definition of Boyen}
Boyen considers two versions of reusable fuzzy extractors, first where the adversary sees $p_1,..., p_q$~(outsider security~\cite[Definition 6]{Boyen2004}) and tries to learn about the values $w_1,..., w_q$ or the keys $r_1,..., r_q$.  Second, where the adversary controls some subset of the servers and can key generation on arbitrary $p_i'$~(insider security~\cite[Definition 7]{Boyen2004}).   This allows the adversary to learn a subset of keys $r_i$~(by performing key generation on the valid $p_i$).  This definition makes sense when servers are compromised~(after enrollment) and act maliciously.  In both definitions, the adversary creates a perturbation function $f_i$ after seeing $p_1,..., p_{i-1}$~(and generated keys for outsider security) and the challenger generates $w_i = f_i(w_1)$.  The definition is parameterized by the class of allowed perturbation functions.

The definition of Boyen considers a single adversary.  \defref{def:outsider fuzz ext} splits the adversary into two parts, one of which is information-theoretic and another that is computationally bounded.  The functions $f_2,..., f_q$ can be thought of as a single adversary that sees all prior state.  However, to provide meaningful security in the computational setting, we cannot have communication between $f_2,..., f_q$ and $D$.  This is because $f_2,..., f_q$ are computationally unbounded and thus could break any security that only holds against computationally bounded adversaries.
Because these two adversaries do not communicate we strengthen the definition by allowing the perturbation functions,  $f_i$, to see the original sample $w_1$.\footnote{An alternative would be to have a single computationally bounded adversary that provides perturbation functions but does not see the original reading $w_1$ as in the definition of Boyen.  \consref{cons:sampling} also satisfies this alternative adaption of Boyen's definition.}  This was not allowed in the definition of Boyen, because the adversary was not split-state, and so giving $w_1$ to the adversary would make security impossible.

%Finally, we allow the distinguishing adversary to see other keys \lnote{we do? then why do we call our definition ``outsider''?}.  This is strictly stronger than the definition of Boyen \lnote{you mean than the outsider security definition of Boyen?}.  However, it is weaker than the notion of insider reusable fuzzy extractors where the adversary sees key generation on arbitrary $p$.  This \lnote{what is ``this''?} is a mixture of reusability and robustness~\cite{dkrs2006} of the fuzzy extractor.  Our reusable construction~(\consref{cons:sampling}) is not insider secure \lnote{according to Boyen's notion, right? Why not say ``not robust'' and/or clarify that $r_i$ can leak but $p_i$ cannot be modified.}.
\subsection{Obfuscation}
Our constructions will use obfuscation for two types of circuits: point functions and digital lockers. The family of point functions $\mathtt{I}_n = \{I_w\}_{w \in \zo^n}$ defined as follows:
\[
I_w(x):\begin{cases} 1 & x=w\\0 & \text{otherwise}\end{cases}.
\]
and the class of digital lockers is $\mathtt{I}_n = \{I_{w, r}\}_{w \in \zo^n, r\in\zo^\kappa}$ defined as follows:
\[
I_{w, r}(x):\begin{cases} r & x=w\\\perp & \text{otherwise}\end{cases}.
\]
The required notion of obfuscation is virtual grey-box (VGB) introduced in \cite{bitansky2010strong}. This notion is weaker then the standard notion of virtual black-box (\cite{barak2001possibility}), as it allows the simulator to run in unbounded time while making at a polynomial number of oracle queries to the function. In addition, we require that the obfuscation is composable and secure with respect to auxiliary input.

Composable auxiliary-input VGB obfuscators for point functions and digital lockers are constructed in \cite[Theorem 6.1]{bitansky2010strong} (following \cite{canetti2008obfuscating,CKVW10}) from the Strong Vector Decision Diffie-Hellman assumption, which is a generalization of the strong DDH assumption of \cite{canetti1997towards} for tuples of points. As described in the Introduction, they can also be constructed by assuming strong properties of cryptographic hash functions~\cite{canetti1997towards}, \cite[Section 4]{lynn2004positive}, \cite[Section 3.2]{canetti2008obfuscating}, \cite[Section 8.2.3]{dakdoukThesis} (such a construction results in an obfuscated function that, at a few hard-to-find-inputs, differs from the original function; this change does not materially affect our results).

@@

\begin{proof} It suffices to argue correctness and security.

\textbf{Correctness:} Correctness follows immediately: for any given $i$, the probability that $v'_i=v_i$ is at least $1/p(n)$. By choosing $\ell$ sufficiently bigger than $1/p(n)$, we guarantee that for at least one $i$,  $v'_i=v_i$ with all but negligible probability
(indeed, because the $\ell$ samples are independent, the probability this does not happen is at most $(1-1/p(n))^\ell \le e^{-\ell/p(n)}$, using $1+x \le e^x$).

\textbf{Security:}
We now argue security.
 Our goal is to show that for all $s_{sec} = \poly(n)$ there exists $\epsilon_{sec} =\ngl(n)$ such that $\delta^{\mathcal{D}_{s_{sec}}}((R, P), (U, P))\le \epsilon_{sec}$. Let $D$ be a distinguisher of size at most $s_{sec}$. We want to bound 
\[
| \expe[D(R, P)] - \expe[D(U, P)]|
\]
by a negligible function.

Adding together Equation~\ref{eq:dist before sampling}, Equation~\ref{eq:sim cannot distinguish samp}, and Equation~\ref{eq:dist before sampling} in which $R$ is replaced with $U$, we obtain the desired result by the triangle inequality.  This completes the proof of \thref{thm:sampling}.
\end{proof}

@@

\subsection{Reusability of \consref{cons:sampling}}
The reusability of \consref{cons:sampling} follows from the security of the VGB obfuscator with auxiliary input. Consider any $q = \poly(n)$ number of reuses.  For each fixed $i\in \{1,..., q\}$, we can treat the keys $r_1,\dots, r_{i-1}, r_{i+1}, \dots, r_q$ and the strings $p_1, \dots, p_q$ as auxiliary input to the obfuscation adversary. The result follows by simulatability of this adversary, using the same argument as the proof of Theorem~\ref{thm:sampling} above. Note that the simulator in the definition of composable obfuscation is required to function for arbitrary circuits in the family even if the choice of these circuits depends on the previous obfuscations, which allows the $i$th reading of $w$ to be chosen depending on public values $p_1,..., p_{i-1}$, as required by definition of reusability.

@@

\noindent
The following theorem shows the security of this construction.

\begin{theorem}
\label{thm:info theory sec}
%Let $n$ be a security parameter and let $\gamma = \omega(\log n)$.
Let $\mathcal{W}$ be a family of $(\alpha = \Omega(1), \beta\leq \gamma(1-\Theta(1)))$-sparse block sources over $\mathcal{Z}^\gamma$ and let $\cond: \mathcal{Z} \times \zo^d\rightarrow \mathcal{Y}$ be a $(\alpha, \tilde{\alpha}, \epsilon_{cond})$-randomness conductor.  Define $\mathcal{V}$ as the family of all distributions with min-entropy at least $\tilde{\alpha}(\gamma-\beta)$ and let $(\gen', \rep')$ be $(\mathcal{Y}^\gamma, \mathcal{V}, \kappa, t, \epsilon_{fext})$-fuzzy extractor with error $\delta$.\footnote{We actually need $(\gen', \rep')$ to be an average case fuzzy extractor~(see \cite[Definition 4]{DBLP:journals/siamcomp/DodisORS08} and the accompanying discussion).  Most known constructions of fuzzy extractors are average-case fuzzy extractors.  For simplicity we refer to $\gen', \rep'$ as simply a fuzzy extractor.}  Then $(\gen, \rep)$ is a $(\mathcal{Z}^\gamma, \mathcal{W}, \kappa, t, \gamma\epsilon_{cond}+\epsilon_{fext})$-fuzzy extractor with error $\delta$.
\end{theorem}
\begin{proof}
Let $W\in \mathcal{W}$.
It suffices to argue correctness and security.  We first argue correctness.

\textbf{Correctness:}
When $w_i = w_i'$, then $\cond(w_i , seed_i) = \cond(w_i', seed_i)$ and thus $v_i = v_i'$.  Thus, for all $w, w'$ where $\dis(w, w')\le t$, then $\dis (v, v')\le t$.  Then by correctness of $(\gen', \rep')$, $\Pr[(r, p)\leftarrow \gen'(v) \wedge r'\leftarrow \rep(v',p) \wedge r' = r]\ge 1-\delta$.

\textbf{Security:}
We now argue security.  Denote by $seed$ the random variable consisting of all $\gamma$ seeds and $V$ the entire string of generated $V_1,..., V_\gamma$.  To show that \[R | P, seed \approx_{\gamma \epsilon_{cond} + \epsilon_{fext}} U | P, seed,\] it suffices to show that $\Hav(V | seed)$ is $\gamma \epsilon_{cond}$ close to a distribution with average min-entropy $\tilde{\alpha}(\gamma - \beta)$.  The lemma then follows by the security of $(\gen', \rep')$.\footnote{Note, again, that $(\gen', \rep')$ must be an average-case fuzzy extractor.  Most known constructions are average-case and we omit this notation.}

\noindent
By a hybrid argument there exists a distribution $Z$ with $\Hav(Z | seed) \ge \tilde{\alpha}(\gamma -\beta)$ where
\[
(\cond(W_\gamma, seed_\gamma), ..., \cond(W_1, seed_1), seed_\gamma,..., seed_1) \approx_{\gamma \epsilon_{cond}} (Z, seed_\gamma,...,  seed_1).\]
%By the security of $(\sketch, \rec)$ we know that $\Hav(Z | seed, ss) \ge \tilde{m}$.  %Note that $U_{\mathcal{Y}}^\gamma$ is independent of $seed_1,..., seed_\gamma$ and so $\Hav(U_{\mathcal{Y}}^\gamma, seed_1,..., seed_\gamma, p)\ge \tilde{m}$.
%That is,
%\[
%(V, seed_1,..., seed_\gamma, p) \approx_{\epsilon\gamma} (U_{\mathcal{Y}}^\gamma, seed_1,..., seed_\gamma, p).\]

This completes the proof of \thref{thm:info theory sec}
\end{proof}

@@

We first define a slightly weaker object than a computational fuzzy extractor: it outputs a key with computational entropy~(instead of a pseudorandom key).  
We call this object a computational fuzzy conductor.  It is the computational analogue of a fuzzy conductor~(introduced by Kanukurthi and Reyzin~\cite{KanukurthiR09}).
Before defining this object, we define conditional computational ``HILL'' (\cite{DBLP:journals/siamcomp/HastadILL99}) entropy.

@@

%\subsection{Reusability of Fuzzy Extractors}
%\label{sec:comparing reusability}

%\subsection{Obfuscation}
%\label{sec:obfuscation def}

@@

%\section{Proof of Security of \consref{cons:sampling}}
%\label{sec:analysis sampling}

\section{Analysis of \consref{cons:first construction}}
\label{sec:construction analysis}

@@
